---
title: "Some Other Exploratory Analysis"
author: "HungTa Chen"
date: "2022-12-06"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Just some other exploratory analysis on the data other than those I did in the cleaning part.

**Read in data**

```{r}
# Read the file
library(stringr)
data = read.csv("./ufc_data.csv")
data = data[1:4727, ]
```

**Remove irrelevant columns**

```{r}
# Drop some irrelevent columns
drop = c("R_fighter", "B_fighter", "Referee", "date", "location", "title_bout")
data_reduced = data[, !(colnames(data) %in% drop)]

# Drop columns related to previous opponents
opp_col = sapply(colnames(data_reduced), str_detect, regex("opp"))
data_reduced = data_reduced[, !(opp_col)]

# Omit rows with NA
data_reduced = na.omit(data_reduced)

data_reduced = data_reduced[, c(2, 3:44, 87, 45:86, 88)]
```

**Change the dimension of the data for further analysis**

```{r}
data_swap = data_reduced[, c(1, 45:87, 2:44)]
combine_data = rbind(data_reduced, data_swap)
combine_data = combine_data[, 1:44]

for(i in 1:ncol(combine_data)){
  if(str_detect(colnames(combine_data)[i], "^B")){
    colnames(combine_data)[i] = paste(strsplit(colnames(combine_data)[i], "_")[[1]][-1], collapse="_")
  }
}

# Drop the fighters who have less than three fights in ufc
newcomer_fights = which((combine_data$wins+combine_data$losses) <= 3)
combine_data = combine_data[(-1)*newcomer_fights, ]
```

As discovered in data cleaning part, the number of strikes attempt (takedown attempt also) is highly related to the number of strikes landed, so we can remove all columns of "attempt".

```{r}
# Drop "strikes attempted" columns 
str_types = c("STR", "HEAD", "BODY", "LEG", "DISTANCE", "CLINCH", "GROUND")
for(i in 1:length(str_types)){
  str = paste(str_types[i], "_att", sep="")
  att_col = sapply(colnames(combine_data), str_detect, regex(str))
  combine_data = combine_data[, !(att_col)]
}

# Drop columns of # of TD landed
td_land_col = sapply(colnames(data_reduced), str_detect, regex("TD_landed"))
data_reduced = data_reduced[, !(td_land_col)]
```

```{r}
# Merge some columns
decision_type = c("Majority", "Split", "Unanimous")
win_by_Decision = numeric(nrow(combine_data))
for(type in decision_type){
  win_by_Decision = win_by_Decision + combine_data[, paste("win_by_Decision_", type, sep = "")]
}
combine_data[, "win_by_Decision"] = win_by_Decision

# Construct some new features
combine_data[, "win_by_Decision_pct"] = combine_data$win_by_Decision/combine_data$wins
combine_data[, "win_by_KO.TKO_pct"] = combine_data$win_by_KO.TKO/combine_data$wins
combine_data[, "win_by_Submission_pct"] = combine_data$win_by_Submission/combine_data$wins
```

**Compare the stats between different weight classes**

```{r}
weight_classes =levels(as.factor(combine_data$weight_class))
bw_stat = combine_data[combine_data$weight_class == "Bantamweight", ]
fw_stat = combine_data[combine_data$weight_class == "Featherweight", ]
flw_stat = combine_data[combine_data$weight_class == "Flyweight", ]
hw_stat = combine_data[combine_data$weight_class == "Heavyweight", ]
lhw_stat = combine_data[combine_data$weight_class == "LightHeavyweight", ]
lw_stat = combine_data[combine_data$weight_class == "Lightweight", ]
mw_stat = combine_data[combine_data$weight_class == "Middleweight", ]
ww_stat = combine_data[combine_data$weight_class == "Welterweight", ]
wbw_stat = combine_data[combine_data$weight_class == "WomenBantamweight", ]
wflw_stat = combine_data[combine_data$weight_class == "WomenFlyweight", ]
wsw_stat = combine_data[combine_data$weight_class == "WomenStrawweight", ]
```

```{r}
for(i in c(2:19, 33, 34, 36, 38:40)){
  boxplot(flw_stat[,i], bw_stat[,i], fw_stat[,i], lw_stat[,i], ww_stat[,i], mw_stat[,i], lhw_stat[,i], hw_stat[,i], main=colnames(fw_stat)[i], names=c("FLW", "BW", "FW", "LW", "WW", "MW", "LHW", "HW"), col=rainbow(8), outline=F)
}
```

From the side-by-side boxplots above, we can see a lot of interesting facts.

1.  The stereotype that Heavyweight fighters barely grapple (wrestling, jiu-jitsu, etc) is real.

    We can see that Heavyweight has the lowest median in almost every stat related to grappling, including "*average takedown percentage*", "*average submission attempt*", "*average reverse*", "*average takedown landed*", "*average clinch strikes landed*", "*average control time*" and "*win by submission percentage*".

    Actually according to the plots, higher weight classes (Middleweight, LightHeavyweight, Heavyweight) generally seem to perform worse in grappling division.

2.  The power of one fighter must have something to do with their weight.

    Firstly the concept that Heavyweight fighters have the most power in their hands is also true. We can see that Heavyweight has the highest "*average knockdown*" among all divisions. Moreover, Heavyweight has the highest "*win by KO/TKO percentage*" among all divisions (also has the lowest "*win by Decision percentage*"), actually it's way higher than other weight classes, which mean Heavyweight fighters frequently end the fight by knocking out their opponents (that's probably why they have the lowest "*average total time fought*" according to the plot).

    In contrast, Flyweight has the lowest "*average knockdown*" and the lowest "*win by KO/TKO percentage*" (highest "*win by decision percentage*" also), which means Flyweight fighters really don't have enough power to hurt their opponent so that their fights also end in judge decision.

    Additionally, we can actually see in the boxplots that higher weight class has higher "*win by KO/TKO percentage*" and lower "*win by decision percentage*", which further suggests the relationship between one's weight and power.

3.  
